---
title: "ADDEP Analysis"
author: "AKV"
date: "3/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
setwd("/Users/user/Documents/AK/ADDEP/")
load("ADDEP.RData")
```

#Analysis for ADDEP dataset 

##Data Cleaning has been done separately. Click [here](https://github.com/AnhKhoaVo/ADDEP/blob/master/ADDEP_Clean_Analysis.R) for the R code.

##Descriptive statistics for Baseline Association 

###Lower Extremity Motor Scores at baseline analyis
```{r, descriptive for LEMS baseline, warning=FALSE }
library(table1)
y <- subset(ADDEP_3, ASIA_LEVEL_ADM==c("C", "T")&!is.na(CRLOWALBUMIN))

#Descriptives stats between included (FALSE) and excluded (TRUE) in LEMS baseline
table1(~SEX+AGE_BINARY+REVIEWASIAGRADEADM+ASIA_LEVEL_ADM+CRLOWALBUMIN|is.na(LOWER_MS_REHAB), data=y)
```

### AIS Grades at baseline analysis 
```{r, descriptive for AIS baseline, warning=FALSE }
#Descriptives stats between included (FALSE) and excluded (TRUE) in AIS Grades baseline
table1(~SEX+AGE_BINARY+REVIEWASIAGRADEADM+ASIA_LEVEL_ADM+CRLOWALBUMIN|is.na(ASIAGRADE_WALK), data=y)
```

#Baseline Association 

These following two figures are to see if Albumin is associated with LEMS and AIS grades at baseline:
```{graphs, warning=FALSE}
library(ggplot2)
library(ggpubr)
cleanup <- theme(panel.grid.major = element_blank(),
                 panel.grid.minor = element_blank(),
                 panel.background = element_blank(),
                 axis.line = element_line(colour="gray"),
                 plot.title = element_text(size = rel(1.2), 
                                           vjust = 1.5, 
                                           hjust = 0.5),
                 plot.subtitle = element_text(size = rel(1), 
                                              vjust = 1.5, 
                                              hjust = 0.5,
                                              face="italic"),
                 panel.border=element_rect(colour="gray",
                                           size=1, 
                                           fill=NA), 
                 panel.spacing = unit(0.2, "lines"),
                 axis.ticks.y = element_line(colour="gray"), 
                 axis.ticks.x = element_line(colour="gray"))

ALB_LEMS <- ggplot(data=subset(ADDEP_3, !is.na(ASIAGRADE_WALK)&ASIA_LEVEL_DIS==c("C", "T")&!is.na(CRLOWALBUMIN)), aes(x=CRLOWALBUMIN, y=LOWER_MS_REHAB, colour = ASIAGRADE_WALK))+
  geom_jitter(width=0.2, alpha=0.6)+
  scale_x_continuous(limits = c(0,5))+
  cleanup+
  theme(axis.text=element_text(size=11),
        axis.title=element_text(size=12),
        strip.text.x = element_text(size = 12),
        panel.border = element_blank(),
        legend.position='bottom',
        legend.key=element_blank())+
  labs(colour='AIS Grades') +
  labs(y="LEMS at Rehab Admission", x="Albumin concentration [g/dL]") 

ALB_AIS <- ggplot(data=subset(ADDEP_3, !is.na(ASIAGRADE_WALK)&ASIA_LEVEL_DIS==c("C", "T")&!is.na(CRLOWALBUMIN)), aes(y=CRLOWALBUMIN, x=ASIAGRADE_WALK, colour=ASIAGRADE_WALK))+
  geom_boxplot()+
  geom_jitter(width=0.2, alpha=0.6)+
  scale_y_continuous(limits = c(0,5))+
  cleanup+
  theme(axis.text=element_text(size=11),
        axis.title=element_text(size=12),
        strip.text.x = element_text(size = 12), 
        axis.title.x=element_blank(),
        axis.ticks.x=element_blank(), 
        axis.text.x=element_blank(), 
        panel.border = element_blank(),
        legend.key=element_blank())+
  labs(colour='AIS Grades')+
  labs(y="Albumin concentration [g/dL]", x="AIS grades") 
```


```{r, both graphs, warning=FALSE, fig.width=14, fig.height=12, message=FALSE}
library("ggpubr")
ggarrange(ALB_LEMS, ALB_AIS, nrow=2, labels = c("A", "B"),common.legend = TRUE, legend = "bottom")
```

#URP for Admission 
For baseline association, we included only C and T participants and excluded all NAs in Albumin. In addition, all NAs in the outcomes (LEMS baseline, and AIS grades baseline) are also excluded. 

##URP for LEMS baseline:

```{r, URP LEMS baseline, warning=FALSE, fig.width=14, fig.height=12, message=FALSE}
library("expss")
library("party")
ADDEP_3 = apply_labels(ADDEP_3, CRLOWALBUMIN = "Min Albumin Concentration", 
                       LOWER_MS_REHAB = "LEMS at Admission", ASIAGRADE_WALK = "AIS Grades at Admission")

URP_LEMS_baseline <-use_labels(ADDEP_3, ctree(LOWER_MS_REHAB ~ CRLOWALBUMIN, controls=ctree_control(testtype = "Bonferroni",
                                                                                                    mincriterion = 0.95), data=subset(..data, !is.na(LOWER_MS_REHAB)&ASIA_LEVEL_ADM==c("C", "T")&!is.na(CRLOWALBUMIN))))

plot(URP_LEMS_baseline)
```

URP for AIS grades baseline:

```{r, URP AIS baseline, warning=FALSE, fig.width=14, fig.height=12, message=FALSE}
library("expss")

URP_AIS_baseline <-use_labels(ADDEP_3, ctree(ASIAGRADE_WALK ~ CRLOWALBUMIN,controls=ctree_control(testtype = "Bonferroni",
                                                                                                  mincriterion = 0.95), data=subset(..data, !is.na(ASIAGRADE_WALK)&ASIA_LEVEL_ADM==c("C", "T")&!is.na(CRLOWALBUMIN))))

plot(URP_AIS_baseline, terminal_panel=node_barplot(URP_AIS_baseline))
```

##Descriptive Statistics for 1-year Analysis

###Change Scores Analysis 
```{r, descriptive Change Scores, warning=FALSE}
#Shared dataset between 2 outcomes for 1 year
z<-subset(ADDEP_3,!(Walk_Admission==1)&ASIA_LEVEL_ADM==c("C", "T")&!is.na(CRLOWALBUMIN))

#Descriptives stats between included (FALSE) and excluded (TRUE) in Change scores
table1(~SEX+AGE_BINARY+REVIEWASIAGRADEADM+ASIA_LEVEL_ADM+CRLOWALBUMIN|is.na(Change_Scores), data=z)
```

###Marked Recovery Analysis
```{r, descriptive Marked Recovery, warning=FALSE}
#Descriptives stats between included (FALSE) and excluded (TRUE) in Marked Recovery 
table1(~SEX+AGE_BINARY+REVIEWASIAGRADEADM+ASIA_LEVEL_ADM+CRLOWALBUMIN|is.na(Marked_Recovery_Annual_2), data=z)
```

#URP for 1 year 

For 1 year, the inclusion criteria are: participants who were non-ambulatory at admission to rehabilitation were included, C & T level only, completed assessment of albumin, and neurological outcomes.

##URP for change scores (univariate)
```{r, URP change scores uni, warning=FALSE, fig.width=14, fig.height=12, message=FALSE}
plot(use_labels(ADDEP_3,ctree(Change_Scores ~ CRLOWALBUMIN, data=subset(..data, !is.na(Change_Scores)&!(Walk_Admission==1)&ASIA_LEVEL_ADM==c("C", "T")&!is.na(CRLOWALBUMIN)))))
```

##Subset data for change scores then split

Splitting data into 70% training, and 30% testing. The evaluation metrics for training sample is:
```{r, URP CS Train, warning=FALSE}
library("caret")
ADDEP_CS_subset <- subset(ADDEP_3, !is.na(Change_Scores)&!(Walk_Admission==1)&ASIA_LEVEL_ADM==c("C", "T")&!is.na(CRLOWALBUMIN))

set.seed(4567)
sample_CS <- ADDEP_CS_subset$Change_Scores %>% 
  createDataPartition(p = 0.7, list = FALSE)
train_ADDEP_CS  <- ADDEP_CS_subset[sample_CS, ]
test_ADDEP_CS <- ADDEP_CS_subset[-sample_CS, ]

URP_CS_1y <- train(
  Change_Scores ~ CRLOWALBUMIN, data = train_ADDEP_CS, method = "ctree", na.action = na.omit, 
  trControl = trainControl("cv", number = 10), tuneLength=.1
)

URP_CS_1y
```

Predicting on the 30% unseeen test data. The evaluation metrics for testing sample is: 
```{r, URP CS Test, warning=FALSE}
predict_CS_1y <- cbind(as.data.frame(predict(URP_CS_1y, newdata=test_ADDEP_CS)), test_ADDEP_CS$Change_Scores)

colnames(predict_CS_1y) <- c("predicted","actuals")
postResample(pred = predict_CS_1y$predicted, obs = predict_CS_1y$actuals)
```

Our model is not overfitting since RMSE between test and train samples are not vastly different. 

##URP for change scores (bivariate)
```{r, URP change scores bi, warning=FALSE, fig.width=14, fig.height=12, message=FALSE}
plot(use_labels(ADDEP_3,ctree(Change_Scores ~ CRLOWALBUMIN+LOWER_MS_REHAB, data=subset(..data, !is.na(Change_Scores)&!(Walk_Admission==1)&ASIA_LEVEL_ADM==c("C", "T")&!is.na(CRLOWALBUMIN)))))
```

##URP for Marked Recovery (univariate)
```{r, URP MR uni, warning=FALSE, fig.width=14, fig.height=12, message=FALSE}
plot(use_labels(ADDEP_3,ctree(Marked_Recovery_Annual_2~CRLOWALBUMIN, data=subset(..data, !is.na(Marked_Recovery_Annual_2)&!(Walk_Admission==1)&ASIA_LEVEL_ADM==c("C", "T")&!is.na(CRLOWALBUMIN)))))
```

#Subset data for Marked Recovery then split

Splitting data is similar to Change Score analysis above. The evaluation metrics for training sample is:
```{r, URP MR Train, warning=FALSE}
ADDEP_MR_subset <- subset(ADDEP_3, !is.na(Marked_Recovery_Annual_2)&!(Walk_Admission==1)&ASIA_LEVEL_ADM==c("C", "T")&!is.na(CRLOWALBUMIN))

set.seed(4567)
sample_MR <- ADDEP_MR_subset$Marked_Recovery_Annual_2 %>% 
  createDataPartition(p = 0.7, list = FALSE)
train_ADDEP_MR <- ADDEP_MR_subset[sample_MR, ]
test_ADDEP_MR <- ADDEP_MR_subset[-sample_MR, ]

URP_MR_1y <- train(
  Marked_Recovery_Annual_2 ~ CRLOWALBUMIN, data = train_ADDEP_MR, method = "ctree", na.action = na.omit, 
  trControl = trainControl("cv", number = 10), tuneLength=.1
)
URP_MR_1y
```

Predicting on the 30% unseeen test data. The AUC for ROC and PR curves are: 
```{r, URP MR Test, warning=FALSE}
library("precrec")
library("ROCR")
predict_MR_1y <- predict(URP_MR_1y, newdata=test_ADDEP_MR, type='prob')
predict_MR_1y <- prediction(predict_MR_1y[,2], test_ADDEP_MR$Marked_Recovery_Annual_2)
predict_MR_1y <- cbind(as.data.frame(predict_MR_1y@predictions), as.data.frame(predict_MR_1y@labels))
colnames(predict_MR_1y) <- c("prob","actuals")
evalmod(scores = predict_MR_1y$prob, labels = predict_MR_1y$actuals)
```

The plot for ROC and PR curves:
```{r, plot ROC and PR curves, warning=FALSE, fig.width=14, fig.height=12, message=FALSE}
autoplot(evalmod(scores = predict_MR_1y$prob, labels = predict_MR_1y$actuals))
```

Confusion matrix: 
```{r, confusion matrix, warning=FALSE}
confusionMatrix(predict(URP_MR_1y, newdata=test_ADDEP_MR), test_ADDEP_MR$Marked_Recovery_Annual_2)
```

Univariate URP for Marked Recovery has 75% accuracy, and it's mostly accurate in predicting for cases who did not achieve marked recovery. 

##URP for Marked Recovery (bivariate)
```{r, URP MR bi, warning=FALSE, fig.width=14, fig.height=12, message=FALSE}
ADDEP_3$REVIEWASIAGRADEADM<- as.factor(ADDEP_3$REVIEWASIAGRADEADM)

plot(use_labels(ADDEP_3,ctree(Marked_Recovery_Annual_2~CRLOWALBUMIN+REVIEWASIAGRADEADM, data=subset(..data, !is.na(Marked_Recovery_Annual_2)&!(Walk_Admission==1)&ASIA_LEVEL_ADM==c("C", "T")&!is.na(CRLOWALBUMIN)))))
```

#Summary 

At admission to rehabiliation, lower serum albumin was associated with more severe spinal cord injury. However, at 1-year post-injury analysis, serum albumin was not significant after adjusting for injury characteristics. This suggest serum albumin is more useful when neurological details at admission (i.e., LEMS or AIS grade at baseline) are not available. 